---
title: "P8105: Homework 2"
author: Alexander Melamed
output: github_document
---

```{r}
library(tidyverse)
```

# Problem 1
The first part of this problem is to import and clean the Mr. Trash Wheel data and to clean it. 
```{r}
trash_df = 
  readxl::read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                     sheet="Mr. Trash Wheel", 
    range ="A2:N406") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls=as.integer(sports_balls)
  )
trash_df
```

The next part of this problem is to read and clean precipitation data for 2017 and 2018. For each, omit rows without precipitation data and add a variable year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}
##Import and clean 2017 & 2018 data
precipitation_2017_df = 
  readxl::read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                     sheet="2017 Precipitation", 
    range ="A2:B14") %>%
  janitor::clean_names() %>% 
  mutate(year=2017)

precipitation_2017_df

##Import and clean 2018 data
precipitation_2018_df = 
  readxl::read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                     sheet="2018 Precipitation", 
    range ="A2:B14") %>%
  janitor::clean_names() %>% 
  mutate(year=2018)

precipitation_2018_df
```